Off the beaten path: \\ Gaia reveals GD-1 stars outside of the main stream

### 2018-07-25 paper/submit-aas/referee-response2.txt

We thank the referee for the suggestions. We believe we have now addressed all of the referee's comments either in the manuscript text or response below. We have also attached a number of figures to the response that were suggested or requested by the referee.

@@

We have attached the requested figure to the report. We do not see correlations between the off-track features described in this paper and the CMD-selected stars with 2-parameter solutions.

We have compared the fraction of good solutions in the spur, in the region between the stream and the spur, and in the stream at the same range of longitudes. We consider the longitude range -35 < phi1 < -30, and 3 ranges in phi2: (-0.4, 0.4), (0.4, 0.9), (0.9, 1.7) for the stream, region between stream and spur, and spur, respectively. In these 3 regions, we find 84%, 88%, 89% of the 5 parameter solution stars pass the unit weight error criteria.

(we have added text to the manuscript, bolded, to describe the points above)

### 2018-07-25 paper/gd1.tex
fix d and add model lines
ref response

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{track_observables.pdf}
\end{center}
\caption{%
Median stream properties computed from the proper motion and CMD filtered GD-1
stars, and the same for a toy model for the stream track.
\textit{Panel (a)}: Sky positions of candidate GD-1 stars.
Lines show a \changes{4th}-order polynomial fit to the median stream track offset by $\pm 0.75^\circ$; the enclosed region is adopted as the nominal stream region.
\changes{\textit{Panels (b) and (c)}: Sky positions of simulated star particles from mock stream models of the GD-1 stream in which the progenitor is currently disrupting (toy model 1, panel b) or the progenitor fully disrupted 500 Myr ago and results in the observed under-density at $\phi_1 \sim -20^\circ$ (toy model 2, panel c).}
\textit{Panel (d)}: Background-subtracted surface-density estimates computed in
successive, overlapping $3^\circ$ windows for both GD-1 data and the model
stream.
\textit{Panel (e)}: Median proper motion along the stream computed in the same
bins as previous panel \changes{(not solar reflex corrected)}.
\textit{Panel (f)}: Same as Panel (d), but for proper motion in stream latitude.
}
\label{fig:track-and-model}
\end{figure}

Clear density variations, also apparent in the 2D positions of the sources, manifest as sharp features in the surface density estimates along the stream.
We have verified that these do not correspond to features in dust extinction nor are they caused by completeness variations in the \gaia\ scanning pattern:
the visible portions of the GD-1 stream are located at high Galactic latitudes ($b > 20^\circ$), and 99\% of our sample has at least 8 \gaia\ visibility periods \changes{and satisfy a selection on the unit weight error (see Equation~C.1 in \citealt{Lindegren:2018})}, promising a robust astrometric solution.
\changes{
Small-scale density variations, however, especially apparent between $-80^\circ < \phi_1 < -50^\circ$, are a consequence of the \gaia\ scanning pattern.
}

>>>>

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{track_observables.pdf}
\end{center}
\caption{%
Median stream properties computed from the proper motion and CMD filtered GD-1
stars, and the same for a toy model for the stream track.
\textit{Panel (a)}: Sky positions of candidate GD-1 stars.
Lines show a \changes{4th}-order polynomial fit to the median stream track offset by $\pm 0.75^\circ$; the enclosed region is adopted as the nominal stream region.
\changes{\textit{Panels (b) and (c)}: Sky positions of simulated star particles from mock stream models of the GD-1 stream in which the progenitor is currently disrupting (toy model 1, panel b) or the progenitor fully disrupted 500 Myr ago and results in the observed under-density at $\phi_1 \sim -20^\circ$ (toy model 2, panel c).}
\textit{Panel (d)}: Background-subtracted surface-density estimates computed in
successive, overlapping $3^\circ$ windows for both GD-1 data and the model
stream.
\textit{Panel (e)}: Median proper motion along the stream computed in the same
bins as previous panel \changes{(not solar reflex corrected)}.
\changes{The two toy models have consistent median proper motions (overlapping dashed lines).}
\textit{Panel (f)}: Same as Panel (d), but for proper motion in stream latitude.
}
\label{fig:track-and-model}
\end{figure}

Clear density variations, also apparent in the 2D positions of the sources, manifest as sharp features in the surface density estimates along the stream.
We have verified that these do not correspond to features in dust extinction nor are they caused by completeness variations in the \gaia\ scanning pattern:
the visible portions of the GD-1 stream are located at high Galactic latitudes ($b > 20^\circ$), and 99\% of our sample has at least 8 \gaia\ visibility periods \changes{and satisfy a selection on the unit weight error (see Equation~C.1 in \citealt{Lindegren:2018})}, promising a robust astrometric solution.
\changes{
In detail, we have compared the fraction of good solutions in the region near the spur, in the region between the stream and the spur, and in the stream at the same range of longitudes.
We consider the longitude range $-35^\circ < \phi_1 < -30^\circ$, and 3 ranges in $\phi_2$: $(-0.4, 0.4)^\circ$, $(0.4, 0.9)^\circ$, $(0.9, 1.7)^\circ$ for the stream, region between stream and spur, and spur, respectively.
In these 3 regions, we find 84\%, 88\%, 89\% of the stars with 5-parameter solutions pass the unit weight error criteria.
Small-scale density variations, however, especially apparent between $-80^\circ < \phi_1 < -50^\circ$, are a consequence of the \gaia\ scanning pattern.
}
